---
title: "Probability and Statistics"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Slides, assignments, and code for my probability and statistics class."
  ) %>%
  meta_name("github-repo" = "djhunter/cryptography") %>%
  meta_viewport() %>%
  meta_social(
    title = "Probability and Statistics",
    url = "https://djhunter.github.io/prob-stat/index.html",
    image = "https://upload.wikimedia.org/wikipedia/commons/1/1f/Pi_statistisch.png",
    image_alt = "Monte Carlo estimation of Pi",
    og_author = c("David J. Hunter")
  )
```

This site provides slides, assignments, and other resources for an upper-division course in probability and statistics. The material in this course draws from the following textbooks.

- [GS]: [Grinstead and Snell's Introduction to Probability](https://math.dartmouth.edu/~prob/prob/prob.pdf), The Chance Project.
- [JWHT]: [An Introduction to Statistical Learning](https://hastie.su.domains/ISLR2/ISLRv2_website.pdf), James, Witten, Hastie, Tibshirani. 
- [ER]: [Probability and Statistics](https://www.utstat.toronto.edu/mikevans/jeffrosenthal/book.pdf), Evans and Rosenthal.

# Slides 

# Assignments

# Other Resources

- [Syllabus](syllabus/ma130hunterS23.pdf) (PDF)
- [Installing R and RStudio](assignments/installRStudio.html)

All of the pages and slides on this site were authored in RMarkdown. The source code is available on GitHub: https://github.com/djhunter/prob-stat

