---
title: "Discrete Probability Distributions"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```


# Overview

For further reading, see [GS], pp. 1-40.

# Simulation of Discrete Probabilities

## Discrete Random Variables

A **random variable** is an expression whose value is the outcome of a random experiment.

- Example: Roll a 6-sided die. Let $X$ be the number that comes up.
    - $P(X = 1) = P(X = 2) = \cdots = P(X = 6) = 1/6$
    - $P(X \leq 4) = 2/3$
- Example: Flip 2 coins. Let $Y$ be the total number of heads.
    - $P(Y = 0) = 0.25$, $P(Y = 1) = 0.5$, $P(Y = 2) = 0.25$

## Distribution Functions

A **distribution function** $m$ assigns a nonnegative number $m(\omega_j)$ to each possible
value $\omega_j$ of a random variable.

- Example: Roll a 6-sided die. Let $X$ be the number that comes up.
    - Possible outcomes: $\omega_1 = 1, \omega_2 = 2, \ldots, \omega_6 = 6$.
    - $m(\omega_j) = 1/6$ for all $j$.
- Example: Flip 2 coins. Let $Y$ be the total number of heads.
    - Possible outcomes: $\omega_1 = 0, \omega_2 = 1, \omega_3 = 2$.
    - $m(\omega_1) = 0.25$, $m(\omega_2) = 0.5$, $m(\omega_3) = 0.25$
    
. . .

Note: A distribution function always has "total mass" 1:

$$
\sum_{\text{all } j} m(\omega_j) = 1
$$

## Example: Peter and Paul

Peter and Paul play a game: flip a coin 40 times. 

- If heads, Peter wins one dollar from Paul. 
- If tails, Peter gives one dollar to Paul.
- Let $X$ be Peter's total winnings. How is $X$ distributed?

(See R Console for a demonstration)

## Simulation using R (summary)

```{r, fig.align='center', fig.width=6, fig.height=3.5}
set.seed(1234)
pwinnings <- replicate(10000, sum(sample(c(-1,1), 40, replace = TRUE)))
probs <- table(pwinnings)/10000
plot(probs)
```

---
#   ## Example: Times in lead
#   
#   - Let $Y$ be the number of times during the game that Peter's winnings are positive.
#   - How is $Y$ distributed?
#   
#   (See R Console)
#   
#   ## Times in lead (summary)
#   
#   ```{r, fig.align='center', fig.width=6, fig.height=3.5}
#   set.seed(1234)
#   pinlead <- replicate(10000, sum(cumsum(sample(c(-1,1), 40, replace = TRUE)) > 0))
#   probs2 <- table(pinlead)/10000
#   plot(probs2)
#   ```
---

---
# Exercise: Random walk (1D) how is distance from zero distributed? what is the average distance from zero?
---

# Discrete Probability Distributions

## Random Variables and Sample Spaces

## Distribution Functions


