---
title: "Important Distributions"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```


# Overview

- Important Discrete Distributions
    - Binomial
    - Geometric
    - Poisson
- Important Continuous Distributions
    - Exponential
    - Functions of a random variable
    - Normal
    - Simulating random draws

For further reading, see [GS], pp. 180-192, 205-215.  See also [ER], pp. 45-47, 54, 57  

# Important Discrete Distributions

## Geometric

A **geometric random variable** $T$ counts the number of trials it takes to obtain a success, if each trial has success probability $p$ and failure probability $q$.

$$
P(T = n) = q^{n-1}p
$$

. . .

Examples: Roll till you get a 6. 

## Binomial

A **binomial random variable** counts the number of successes $X$ in $n$ trials, where each trial has a probability $p$ of success and $q = 1-p$ is the probability of failure.

$$
P(X = k) = b(n,p,k) = \binom{n}{k} p^k q^{n-k}
$$

. . . 

Examples: Peter and Paul, Flip 3 coins

## Poisson

- Suppose, in the binomial distribution, $n$ is very large, and $p$ is very small.
    - e.g., $n$ is the number of houses in a city, $p$ is the probability that the house will burn down on a given day.
    - We might not even know $n$ and $p$, but we know, on average, how many houses burn down in a city each day.
        - $\lambda = np$
    - The **Poisson** distribution approximates a binomial distribution in this situation.

## Poisson distribution function

- Start with binomial: $P(X = k) = \binom{n}{k} p^k (1-p)^{n-k}$
- Substitute $\lambda/n$ for $p$.
- Let $n \rightarrow \infty$, while $k$ is fixed.

## Start with binomial: 

Substitute $\lambda/n$ for $p$.

$$
\begin{align}
P(X = k) &= \binom{n}{k} \left( \frac{\lambda}{n}\right)^k\left( 1- \frac{\lambda}{n}\right)^{n-k} \\
&= \frac{n(n-1)\cdots (n-k+1)}{k!} \left( \frac{\lambda}{n}\right)^k\left( 1- \frac{\lambda}{n}\right)^{n-k} \\
\end{align}
$$

## Take limits

Send $n \rightarrow \infty$. Note $k$ is fixed.

$$
\frac{n(n-1)\cdots (n-k+1)}{n^k} \rightarrow 1
$$

. . . 

$$
\left( 1- \frac{\lambda}{n}\right)^{n-k} = \left( 1- \frac{\lambda}{n}\right)^{n}\left( 1- \frac{\lambda}{n}\right)^{-k} \rightarrow e^{-\lambda} \cdot 1 = e^{-\lambda}
$$

. . . 

We obtain the **Poisson distribution:** $P(X = k) = \displaystyle{\frac{\lambda^k}{k!}e^{-\lambda}}$

## Activity: Binomial and Poisson in R

1. Make a function `bin(n,p,k)` implementing the binomial.
2. Make a function `poi(lambda,k)` implementing the Poisson.
3. Compare.

## Activity Solution

```{r, eval=FALSE}
bin <- function(n,p,k) {
  choose(n,k)*p^k*(1-p)^(n-k)
  }
bin(100, 0.1, 5)
poi <- function(lambda, k) {
  lambda^k/factorial(k)*exp(-lambda)
  }
poi(10, 5)
plot(bin(100, 0.1, 0:100))
plot(poi(10, 0:100))
```

# Important Continuous Distributions

## Exponential

Recall the exponential density function.

$$
f(x) = \begin{cases}
\lambda e^{-\lambda x} & \text{if } x \geq 0 \\
0 & \text{otherwise}
\end{cases}
$$

The constant $\lambda$ is called a **parameter**.

. . .

It is easy to derive the corresponding cumulative distribution function:

$$
F(x) = \int_{-\infty}^x f(t)\, dt = 
\begin{cases}
0 & \text{if } x < 0 \\
1 - e^{-\lambda x} & \text{if } x\geq 0
\end{cases}
$$

## Standard Normal

The **standard normal** density function is the following.

$$
f(x) = \frac{1}{\sqrt{2\pi}} e^{-x^2/2}
$$

. . .

- There is no antiderivative of this function in terms of elementary functions.
- We have to resort to numerical integration to compute probabilities.

## Activity: Normal probabilites by hand

1. Define a function `f(x)` implementing the Normal PDF.
2. Use `integrate` to compute some areas.
3. Compare to the built-in `pnorm` function.

## Activity Solution

```{r, eval=FALSE}
f <- function(x) {1/sqrt(2*pi)*exp(-x^2/2)}
integrate(f, -1, 1)
pnorm(1.645)
integrate(f, -Inf, 1.645)
```

## Functions of a random variable

**Theorem 5.1.** Suppose $X$ is a continuous random variable, and $\phi(x)$ is a strictly increasing function of the range of $X$. Suppose that $X$ and $Y$ have CDFs $F_X$ and $F_Y$. Then
$$
F_Y(y) = F_X(\phi^{-1}(y)).
$$
(proof, example)

. . . 

**Corollary.** $\displaystyle{f_Y(y) = f_X(\phi^{-1}(y))\frac{d}{dy}\phi^{-1}(y)}$

## Normal Distribution with parameters

- Let $X$ be a standard normal random variable, and let $Y = \sigma X + \mu$.
- We have $X = (Y - \mu)/\sigma$, so $\phi^{-1}(y) = (y - \mu)/\sigma$. 
- Use the Corollary to get a Normal distribution with parameters $\mu$ and $\sigma$.

. . .

$$
f_Y(y) = \frac{1}{\sigma\sqrt{2\pi}} \exp\left(\frac{-(y- \mu)^2}{2\sigma^2}\right)
$$


## Simulating random draws

TODO

## Activity: Exponential draws by hand

1. Define a function `Finv(y)` implementing the *inverse* of the exponential CDF (use $\lambda = 1$). 
2. Use your function, and `runif`, to produce some random draws from the exponential distribution.
3. Compare to the built-in `rexp` function.

## Activity Solution

```{r, eval=FALSE}
Finf <- function(y) {
  -log(1-y)
}
hist(rexp(10000))
hist(Finf(runif(10000)))
```

