---
title: "The Central Limit Theorem"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```


# Overview

- Motivating Examples
- Statement of the CLT
- Applications
    - Confidence Intervals
    - The t-Distribution

For further reading, see [GS], pp. 356-361

## The Sample Mean

>- A random sample of $n$ observations corresponds to *independent, identically-distributed* (iid) random variables $X_1, X_2, \ldots X_n$. 
>- Suppose $E(X_i) = \mu$ and $V(X_i) = \sigma^2$ for all $i$.
>- The **sample mean** is then a random variable $\bar{X} = (X_1 + X_2 + \cdots + X_n)/n$.
>    - $E(\bar{X}) = \mu$
>    - $V(\bar{X}) = \sigma^2/n$

# Motivating Examples

## Sample from uniform distribution

- Create sample means for different sample sizes using `runif`.
- Plot these together.
- What are they approaching as $n\rightarrow \infty$?

(Do some R carpentry with tidyverse, ggplot.)

## Sample Solution

```{r, eval=FALSE}
library(tidyverse)
repSize = 10000
simPts <- tibble()
for(n in c(1, 2, 5, 10, 20)) {
  simPts <- simPts %>%
    bind_rows(
      tibble(xbar = replicate(repSize, mean(runif(n))), 
             sample_size = factor(rep(n, repSize)))
    )
}
ggplot(simPts, aes(x = xbar, y = sample_size)) +
  geom_boxplot()

ggplot(simPts, aes(x = xbar, color = sample_size)) +
  geom_boxplot()

ggplot(simPts, aes(x = xbar, color = sample_size)) +
  geom_histogram()

ggplot(simPts, aes(x = xbar, color = sample_size)) +
  geom_histogram(position = "dodge")

ggplot(simPts, aes(x = xbar, color = sample_size)) +
  geom_density()

simPts %>%
  filter(sample_size == 20) %>%
  ggplot(aes(x = xbar)) +
    geom_density()

ss <- 20
simPts %>%
  filter(sample_size == ss) %>%
  ggplot(aes(x = xbar)) +
    geom_density() +
    stat_function(fun = dnorm, 
                  args = list(mean = 0.5, 
                              sd = sqrt(1/12/ss)), 
                  color = "green")
```


# The Central Limit Theorem

## Standardization

Let $X$ be a random variable with mean $\mu$ and variance $\sigma^2$. Then its **standardized** version is
$$
X^*  = \frac{X - \mu}{\sigma}
$$

- $X^*$ has mean 0 and variance 1 (exercise).
- If $\bar{X}$ is the mean of $n$ iid random variables with mean $\mu$ and variance $\sigma^2$, then
$$
\bar{X}^* = \frac{\bar{X} - \mu}{\sigma/\sqrt{n}}
$$
    - Note: $\bar{X}^*$ is the same as $S_n^*$.
    
## Central Limit Theorem

**Theorem 9.6** (Central Limit Theorem). Let $X_1, X_2, \ldots, X_n$ be independent, identically-distributed random variables with mean $\mu$ and variance $\sigma^2$. Then
$$
\lim_{n \rightarrow \infty} P(a < \bar{X}^* < b) = \frac{1}{\sqrt{2\pi}}\int_a^b e^{-x^2/2} \, dx.
$$
In other words, the sample mean is asymptotically normally distributed.

# Applications

## Exercise 10 revisited

Suppose test scores have a mean of 70 and variance 25. Using Chebyshev, we estimated that for a class of size 100,
$$
P(|\bar{X} - 70| \leq 5) \geq 1 - 1/10^2 = 0.99
$$

. . .

Applying the CLT, we have $\bar{X}^*$ is approximately standard normal. 

$$
P(|\bar{X} - 70| \leq 5) = P\left(\frac{|\bar{X} - 70|}{5/\sqrt{100}} \leq \frac{5}{5/\sqrt{100}}\right) = P(|\bar{X}^*| \leq 10)
$$

. . .

```{r}
pnorm(10)-pnorm(-10)
```

## Find the sample size (Exercise 10 revisited)

What class size $n$ would ensure that $P(|\bar{X} - 70| \leq 5) \approx 0.95$? 

- Note: $\sigma = 5/\sqrt{n}$
- Chebyshev: $1 - 1/n = 0.95$, so $n \approx 1/0.05 = 20$.
- CLT: $P(|\bar{X}^*| \leq \sqrt{n}) = 0.95$
    - We want a $z$-score for a tail probability of 0.025.

. . .

```{r}
qnorm(0.025)
```

So $n \approx 4$.

## Confidence Interval (Exercise 10 revisited)

Now suppose the class size $n = 100$. For what $\epsilon$ will $P(|\bar{X} - 70| \leq \epsilon) = 0.99$?

. . .

$$
\begin{align}
P(|\bar{X} - 70| \leq \epsilon) &= P\left(\frac{|\bar{X} - 70|}{5/\sqrt{100}} \leq \frac{\epsilon}{5/\sqrt{30}}\right) = P\left(\bar{X}^* \leq \frac{\epsilon}{5/\sqrt{100}}\right)
\end{align}
$$

. . .

```{r}
qnorm(0.005)
```

$$
\epsilon \approx 2.58 \cdot 5/\sqrt{100} \approx 1.3
$$

The interval $70 \pm 1.3$ is called a **99% confidence interval** for the class average.

# Practical Considerations

## Unknown variance

## The t-Distribution

