---
title: "KNN Classification"
output:
  html_document:
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Please submit your written work as a PDF produced by your method of choice, or as a Word Document or Google Doc. Please also upload a separate R script (`filename.R`) with the code that you used to answer the questions (code only, no output): due 2/21, 11:59pm. 

1. Do Exercises 2 and 7 on pages 52-54 of [JWHT].

2. Install the `palmerpenguins` library:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

Load the library and inspect the `penguins` data frame.

```{r, eval = FALSE}
library(palmerpenguins)
library(tidyverse)
glimpse(penguins)
```

The goal is to predict the categorical variable `species` from the four numeric explanatory variables `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`.

- **Create a plot** that incorporates all five of these variables. Include this plot in your written document, along with a sentence or two speculating on which predictor you think is going to be most effective in predicting `species`.
- **Train and test a KNN model** for predicting `species` from these four numeric variables. You will have to eliminate the rows in `penguins` with missing values (see `?drop_na`), because the `knn` command from the `class` library will not tolerate `NA`s. Choose an appropriately-sized train/test split, and experiment with some different values of `k` to find a model with minimal testing set error rate. 
- **Use your model to predict** the species of the five penguins described in the following data frame. Try a few different values of `k`. Do different values of `k` cause the predicted species to change?

```{r, eval = FALSE}
penToPredict <- tibble(
  bill_length_mm = c(49, 35, 43, 51, 33),
  bill_depth_mm = c(18, 15, 17, 21, 14),
  flipper_length_mm = c(210, 190, 201, 180, 229),
  body_mass_g = c(4560, 3520, 4300, 2850, 6215)
)
```

Summarize your findings in your written document, and also upload your R script as a `.R` file.

